---
output: 
  rmarkdown::github_document:
    html_preview: true
    fig_height: 6
    fig_width: 6
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

**Treemapify provides ggplot2 geoms for drawing [treemaps](http://en.wikipedia.org/wiki/Treemap)**

# Install

Install `ggplot2`, `reshape2`, `plyr`, `devtools` and `ggfittext` if you don't
have them already. `ggfittext` and `treemapify` are installed from github.

```{r eval = FALSE}
install.packages("ggplot2")
install.packages("reshape2")
install.packages("plyr")
install.packages("devtools")
library(devtools)
install_github("wilkox/ggfittext")
install_github("wilkox/treemapify")
```

# Walkthrough

For this example, we'll plot some data on the G-20 group of major world
economies. `treemapify` includes this in the `G20` data frame:

```{r message = FALSE}
library(treemapify)
kable(G20)
```

A treemap has a tile for each observation, with the area of the tile
proportional to a variable. Let's start by drawing a treemap with each tile
representing a G-20 country. The area of the tile will be mapped to the
country's GDP, and the tile's fill colour mapped to its HDI (Human Development
Index). `geom_treemap` is the basic geom for this purpose.

```{r basic_treemap}
ggplot(G20, aes(area = GDP.mil.USD, fill = HDI)) +
  geom_treemap()
```

This plot isn't very useful without the knowing what country is represented by
each tile. `geom_treemap_text` can be used to add a text label to each tile. It
uses the [ggfittext](https://github/wilkox/ggfittext) package to resize the text
so it fits the tile. In addition to standard text formatting aesthetics you
would use in `geom_text`, like ‘fontface’ or ‘colour’, we can pass additional
options specific for `ggfittext`. For example, we can place the text in the
centre of the tile with `place = "centre"`, and expand it to fill as much of the
tile as possible with `grow = TRUE`.

```{r geom_treemap_text}
ggplot(G20, aes(area = GDP.mil.USD, fill = HDI, label = Country)) +
  geom_treemap() +
  geom_treemap_text(
    fontface = "italic",
    colour = "white",
    place = "centre",
    grow = TRUE
  )
```

`geom_treemap` supports subgrouping of tiles within a treemap by passing a
‘subgroup’ aesthetic. Let's subgroup the countries by region, draw a border
around each subgroup with `geom_treemap_subgroup_border`, and label each
subgroup with `geom_treemap_subgroup_text`. As with `geom_treemap_text`,
`geom_treemap_subgroup_text` can take `ggfittext` parameters for text placement
and sizing.

```{r subgrouped treemap}
ggplot(G20, aes(
  area = GDP.mil.USD,
  fill = HDI,
  label = Country,
  subgroup = Region,
  )) +
  geom_treemap() +
  geom_treemap_subgroup_border() +
  geom_treemap_subgroup_text(
    place = "centre",
    grow = T,
    alpha = 0.5,
    colour = "black",
    fontface = "italic",
    min.size = 0
  ) +
  geom_treemap_text(
    colour = "white",
    place = "topleft",
    reflow = T
  )
```

Note that ‘Argentina’ has been hidden. `geom_treemap_text` will hide text labels
that cannot fit a tile without being shrunk below a minimum size, by default 4
points. This can be adjusted with the ‘min.size’ option.

Like any ggplot2 plot, treemapify plots can be faceted, scaled, themed, etc.

```{r complex treemap}
ggplot(G20, aes(area = GDP.mil.USD, fill = Region, label = Country)) +
  geom_treemap() +
  geom_treemap_text(grow = T, reflow = T, colour = "black") +
  facet_wrap( ~ Econ.classification) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  labs(
    title = "The G-20 major economies",
    caption = "The area of each country is proportional to its relative GDP
               within the economic group (advanced or developing)",
    fill = "Region"
  )
```

# Credit

`treemapify` uses the [Squarified Treemap](http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.36.6685) algorithm of Mark Bruls, Kees Huizing and Jarke van Wijk.

The G20 dataset is from the Wikipedia article [G-20 major economies](http://en.wikipedia.org/wiki/G-20_major_economies), which is released under the Creative Commons Attribution-Share-Alike License 3.0.
